A mechanization of semi-simplicial and semi-cubical sets in the calculus of inductive constructions

### 2022-09-24 paper/paper.tex
paper: port over to using bibtex

In the context of typed functional programming, Reynolds' parametricity~\cite{Reynolds72} interprets types as relations characterizing the observational equality of programs of this type. Parametricity can be iterated, and it has been noted that iterated Reynolds' parametricity has a cubical flavor~\cite{Johann17,Altenkirch15,Moulin16}. We obtain a \emph{unary} variant of Reynolds' \emph{binary} parametricity by using predicates instead of relations, in which case we obtain a form of (Kreisel) realizability~\cite{Bernardy12,Moulin16}. It has then been noted that iterated unary parametricity has an augmented simplicial flavor~\footnote{Observation from Hugo Moeneclaey}.

@@

\begin{table}[H]
  \begin{tabularx}{\linewidth}{p{.3\linewidth}|p{.3\linewidth}|p{.3\linewidth}}
    \toprule
            & Set theory     & Type theory \\
    \graymidrule
    Fibered & \cite{Cohen16} &             \\
    \graymidrule
    Indexed &                & Our work    \\
    \bottomrule
  \end{tabularx}
\end{table}

@@

{\scriptsize
\begin{equation*}
  \begin{array}{l}
    \begin{array}{lll}
      \mbox{\textit{squares}}     \\
      [3mm]\mbox{\textit{lines}}  \\
      [3mm]\mbox{\textit{points}} \\
      [1mm]\end{array}
    \framebox{$\begin{array}{lll}
                     \overbrace{\raisebox{0.05cm}{\framebox{$
                           \begin{array}{ll}
                  \raisebox{0cm}{\hspace{.5cm}{\layer[1][0]}}              & \framebox{$a_{0\unitpoint}$} \\
                  [1mm]\overbrace{\framebox{$a_{00}$}~\framebox{$a_{01}$}} &
                  \raisebox{-0.1cm}{\hspace{0.1cm}\filler[1][0]\!\!\!\!\!\!\!\!}
                  \\
                \end{array}
                         $}}~
                     {\raisebox{0.05cm}{\framebox{$
                           \begin{array}{ll}
                  \raisebox{0cm}{\hspace{.5cm}{\layer[1][0]}}              & \framebox{$a_{1\unitpoint}$} \\
                  [1mm]\overbrace{\framebox{$a_{10}$}~\framebox{$a_{11}$}} &
                  \raisebox{-0.1cm}{\filler[1][0]\!\!\!\!\!\!\!\!}
                \end{array}
                         $}}}}^{\layer[2][0]}~\stackrel{\raisebox{9.5mm}{$\framebox{$
                               \begin{array}{ll}
                      \raisebox{0cm}{\hspace{.5cm}{$\layer[2][1]$}}                                & \framebox{$a_{\unitpoint\unitpoint}$} \\
                      [1mm]\overbrace{\framebox{$a_{\unitpoint 0}$}~\framebox{$a_{\unitpoint 1}$}} &
                      \raisebox{-0.1cm}{ \hspace{0.1cm}\filler[2][1]\!\!\!\!\!\!\!\!}
                    \end{array}$}$}}
                     {\raisebox{-0.4cm}{\hspace{2.5cm}\filler[2][0]\!\!\!\!\!\!\!}}
                   \end{array}$}
    ~~~\parbox{2cm}{{\it additionally, each atomic component at dimension $n$ is a \filler[n][n]}} \\\\
    [-1mm]\qquad\qquad\quad~~
    \begin{tikzcd}[column sep=5.2em, row sep=1em]
      \phantom{\bullet}\arrow[r, leftrightarrow] & \phantom{\bullet}
    \end{tikzcd}
    \mbox{\framep[1][1]}
    \qquad
    \begin{tikzcd}[column sep=5.2em, row sep=1em]
      \phantom{\bullet}\arrow[r, leftrightarrow] & \phantom{\bullet}
    \end{tikzcd}
    \mbox{\framep[1][1]}                                                                           \\
    \qquad\qquad~~
    \begin{tikzcd}[column sep=32em, row sep=1em]
      \phantom{\bullet}\arrow[r, leftrightarrow] & \phantom{\bullet}
    \end{tikzcd}
    \mbox{\framep[2][1]}                                                                           \\
    \qquad\qquad~~
    \begin{tikzcd}[column sep=39em, row sep=1em]
      \phantom{\bullet}\arrow[r, leftrightarrow] & \phantom{\bullet}
    \end{tikzcd}
    \mbox{\framep[2][2]}                                                                           \\
  \end{array}
\end{equation*}
}

@@

The proof of $\cohf$ require a higher-dimensional coherence condition which we obtain by working in \texttt{HSet}. If the proofs of the same equality were not equated, there would be a need for arbitrary many higher-dimensional coherences (see e.g.~\cite{Herbelin15} for a discussion on the de facto need for recursive higher-dimensional coherence conditions in formulating higher-dimensional structures in type theory).

@@

\newpage
\begin{thebibliography}{10}
  \bibitem[Reynolds72]{Reynolds72}
  Reynolds, J. C. (1972, August). Definitional interpreters for higher-order programming languages. \emph{In Proceedings of the ACM annual conference-Volume 2} (pp. 717-740).

  \bibitem[Johann17]{Johann17}
  Johann, P., \& Sojakova, K. (2017). Cubical categories for higher-dimensional parametricity. \emph{arXiv preprint arXiv:1701.06244}.

  \bibitem[BernardyMoulin]{Bernardy12}
  Bernardy, J. P., \& Moulin, G. (2012, June). A computational interpretation of parametricity. \emph{In 2012 27th Annual IEEE Symposium on Logic in Computer Science} (pp. 135-144). IEEE.

  \bibitem[AltenkirchKaposi]{Altenkirch15}
  Altenkirch, Thorsten, and Ambrus Kaposi. "A nominal syntax for internal parametricity." \emph{TYPES 2015} (2015): 21.

  \bibitem[Moulin16]{Moulin16}
  Moulin, G. (2016). \emph{Internalizing parametricity.} Department of Computer Science and Engineering, Chalmers University of Technology.

  \bibitem[Bez]{Bezem14}
  Bezem, M., Coquand, T., \& Huber, S. (2014, July). A model of type theory in cubical sets. \emph{In 19th International Conference on Types for Proofs and Programs (TYPES 2013)} (Vol. 26, pp. 107-128). Wadern, Germany: Schloss Dagstuhl–Leibniz Zentrum fuer Informatik.

  \bibitem[CubAgda]{Vezzosi21}
  Vezzosi, A., Mörtberg, A., \& Abel, A. (2021). Cubical Agda: a dependently typed programming language with univalence and higher inductive types. \emph{Journal of Functional Programming, 31}.

  \bibitem[Takeuti–Gandy]{Barras15}
  Barras, B., Coquand, T., \& Huber, S. (2015). A generalization of the Takeuti–Gandy interpretation. \emph{Mathematical structures in computer science}, 25(5), 1071-1099.

  \bibitem[CCHM]{Cohen16}
  Cohen, C., Coquand, T., Huber, S., \& Mörtberg, A. (2016). Cubical type theory: a constructive interpretation of the univalence axiom. \emph{arXiv preprint arXiv:1611.02108}.

  \bibitem[Angiuli]{Angiuli17}
  Carlo Angiuli, Guillaume Brunerie, Thierry Coquand, Kuen-Bang Hou (Favonia), Robert Harper, \& Daniel R. Licata. \emph{arXiv preprint}.

  \bibitem[Fri]{Friedman08}
  Friedman, G. (2008). An elementary illustrated introduction to simplicial sets. \emph{arXiv preprint arXiv:0809.4221}.

  \bibitem[Rie]{Riehl11}
  Riehl, E. (2011). A leisurely introduction to simplicial sets. \emph{Unpublished expository article available online from the author's web page}.

  \bibitem[Her]{Herbelin15}
  Herbelin, H. (2015). A dependently-typed construction of semi-simplicial types. \emph{Mathematical Structures in Computer Science, 25}(5), 1116-1131.

  \bibitem[CoqInCoq]{Barras97}
  Barras, B., \& Werner, B. (1997). Coq in coq. \emph{Available on the WWW.}
\end{thebibliography}

>>>>

In the context of typed functional programming, Reynolds' parametricity~\cite{reynolds72} interprets types as relations characterizing the observational equality of programs of this type. Parametricity can be iterated, and it has been noted that iterated Reynolds' parametricity has a cubical flavor~\cite{johann17,altenkirch14,moulin16}. We obtain a \emph{unary} variant of Reynolds' \emph{binary} parametricity by using predicates instead of relations, in which case we obtain a form of (Kreisel) realizability~\cite{bernardy12,moulin16}. It has then been noted that iterated unary parametricity has an augmented simplicial flavor~\footnote{Observation from Hugo Moeneclaey}.

@@

\begin{table}[H]
  \begin{tabularx}{\linewidth}{p{.3\linewidth}|p{.3\linewidth}|p{.3\linewidth}}
    \toprule
            & Set theory  & Type theory \\
    \graymidrule
    Fibered & \cite{cchm} &             \\
    \graymidrule
    Indexed &             & Our work    \\
    \bottomrule
  \end{tabularx}
\end{table}

@@

{\scriptsize
\begin{equation*}
  \begin{array}{l}
    \begin{array}{lll}
      \mbox{\textit{squares}}     \\
      [3mm]\mbox{\textit{lines}}  \\
      [3mm]\mbox{\textit{points}} \\
      [1mm]\end{array}
    \framebox{$\begin{array}{lll}
                     \overbrace{\raisebox{0.05cm}{\framebox{$
                           \begin{array}{ll}
                  \raisebox{0cm}{\hspace{.5cm}{\layer[1][0]}}              & \framebox{$a_{0\unitpoint}$} \\
                  [1mm]\overbrace{\framebox{$a_{00}$}~\framebox{$a_{01}$}} &
                  \raisebox{-0.1cm}{\hspace{0.1cm}\filler[1][0]\!\!\!\!\!\!\!\!}
                  \\
                \end{array}
                         $}}~
                     {\raisebox{0.05cm}{\framebox{$
                           \begin{array}{ll}
                  \raisebox{0cm}{\hspace{.5cm}{\layer[1][0]}}              & \framebox{$a_{1\unitpoint}$} \\
                  [1mm]\overbrace{\framebox{$a_{10}$}~\framebox{$a_{11}$}} &
                  \raisebox{-0.1cm}{\filler[1][0]\!\!\!\!\!\!\!\!}
                \end{array}
                         $}}}}^{\layer[2][0]}~\stackrel{\raisebox{9.5mm}{$\framebox{$
                               \begin{array}{ll}
                      \raisebox{0cm}{\hspace{.5cm}{$\layer[2][1]$}}                                & \framebox{$a_{\unitpoint\unitpoint}$} \\
                      [1mm]\overbrace{\framebox{$a_{\unitpoint 0}$}~\framebox{$a_{\unitpoint 1}$}} &
                      \raisebox{-0.1cm}{ \hspace{0.1cm}\filler[2][1]\!\!\!\!\!\!\!\!}
                    \end{array}$}$}}
                     {\raisebox{-0.4cm}{\hspace{2.5cm}\filler[2][0]\!\!\!\!\!\!\!}}
                   \end{array}$}
    ~~~\parbox{2cm}{additionally, each atomic component at dimension $n$ is a \filler[n][n]} \\\\
    [-1mm]\qquad\qquad\quad~~
    \begin{tikzcd}[column sep=5.2em, row sep=1em]
      \phantom{\bullet}\arrow[r, leftrightarrow] & \phantom{\bullet}
    \end{tikzcd}
    \mbox{\framep[1][1]}
    \qquad
    \begin{tikzcd}[column sep=5.2em, row sep=1em]
      \phantom{\bullet}\arrow[r, leftrightarrow] & \phantom{\bullet}
    \end{tikzcd}
    \mbox{\framep[1][1]}                                                                     \\
    \qquad\qquad~~
    \begin{tikzcd}[column sep=32em, row sep=1em]
      \phantom{\bullet}\arrow[r, leftrightarrow] & \phantom{\bullet}
    \end{tikzcd}
    \mbox{\framep[2][1]}                                                                     \\
    \qquad\qquad~~
    \begin{tikzcd}[column sep=39em, row sep=1em]
      \phantom{\bullet}\arrow[r, leftrightarrow] & \phantom{\bullet}
    \end{tikzcd}
    \mbox{\framep[2][2]}                                                                     \\
  \end{array}
\end{equation*}
}

@@

The proof of $\cohf$ require a higher-dimensional coherence condition which we obtain by working in \texttt{HSet}. If the proofs of the same equality were not equated, there would be a need for arbitrary many higher-dimensional coherences (see e.g.~\cite{herbelin15} for a discussion on the de facto need for recursive higher-dimensional coherence conditions in formulating higher-dimensional structures in type theory).

@@

\bibliographystyle{alpha}
\bibliography{paper}
